---
title: "module3"
author: "Sushil Paudel"
date: "2024-02-13"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{module3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Vignette tasks



```{r, error = TRUE}
set.seed(1)
v1 <- sample(as.integer(1:100), 20, replace = TRUE)

v2 <- sample(as.integer(1:10), 10, replace = TRUE)

v3 <- runif(20, min = 0, max = 1)

v4 <- sample(LETTERS[v2], 20, replace = TRUE)

print(v1)
print(v2)
print(v3)
print(v4)



```


### Create a matrix **m1** from **v1** and **v2**
```{r, error = TRUE}
m1 <- cbind(v1, v2)
m1

m2 <- cbind(v2, v3)
m2
```


### Create a data.frame named DF by binding **v1**, **v2**, **v3**, **v4**.
```{r, error = TRUE}
DF <- data.frame(v1, v2, v3, v4)
DF
```

###**Question:- Why shouldn’t you use the name “df” to name this object?**
It is because "df" is reserve name in this package.



### Change the column names of DF to “a”, “b”, “c”, and “grp”.
```{r,  error = TRUE}
DF <- data.frame(a = v1, b = v2, c = v3, grp = v4)
DF
```

### **Question: why is it okay to rename a column in DF “c”, but not to rename v3 as c?**
I am not understanding this question. Since we are changing v3 to c which is the column in DF.
Is it some particular question that was discussed in class and might be I missed?

### Look at the str of DF. What kind of data type is “grp”?
  **Answer = the data type of "grp" is character.**  
  
    How would you change it between character and factor and back again?

```{r,  error = TRUE}
str(DF)

DF$grp <- as.factor(DF$grp)
str(DF)

DF$grp <- as.character(DF$grp)
str(DF)

```


### Demonstrate the use of your ner function:
    
    . Use my_calc to sum and divide the following pairs of objects

```{r,  error = TRUE}
library(sus346)
s1 <- my_calc(v1, v2, "sum")
s1
d1 <- my_calc(v1, v2, "divide")
d1

s2 <- my_calc(v2, v3, "sum")
s2
d2 <- my_calc(v2, v3, "divide")
d2

s3 <- my_calc(v1, v3, "sum")
s3
d3 <- my_calc(v1, v3, "divide")
d3

s4 <- my_calc(m1, m1, "sum")
s4
d4 <- my_calc(m1, m1, "divide")
d4

s5 <- my_calc(m1, m2, "sum")
s5
d5 <- my_calc(m1, m2, "divide")
d5
```

### Using my_calc with DF:
    
    Multiply row1, column1 by row1, column2
    Multiply rows 1-10, column 1-3 by rows1-10, columns 1-3
    Divide the values in columns “a” and “b” by the values in column “c” of DF corresponding to the rows in “grp” that have values “E”. 
    In selecting the columns, use their names, rather than their index numbers, to select them

```{r,  error = TRUE}
my_calc(DF[1, 1], DF[1, 2], "multiply")

my_calc(DF[1:10, 1:3], DF[1:10, 1:3], "multiply")

o <- DF[DF$grp == "E", ]

my_calc(o$a, o$c, "divide")

my_calc(o$b, o$c, "divide")

```
### Repeat the last of the aforementioned calculations, but this time do not use my_calc.
      
      Simply do it using the two DF subsets with the / operator

```{r,  error = TRUE}
o$a / o$c

o$b / o$c
```
### Demonstrate the use of *apply:

    Combine v1, m1, and m2 into a list l
    Write an lapply that applies sum to l. This one does not need an anonymous function
    Now write an lapply that iterates over the elements of l, and uses your my_calc to multiply each element of l by v2
    
  **Hints:**
    *You will need to use an anonymous function*
    *The interior of your function (i.e. the part between ({}) will use the my_calc function.*
    
    Do the same but with sapply

```{r,  error = TRUE}
l <- list(v1, m1, m2)
l

lapply(1:length(l),  function(x) sum(l[[x]]))

lapply(1:length(l), function(x){
  my_calc(l[[x]], v2, "multiply")
})

sapply(1:length(l),  function(x) sum(l[[x]]))

```


