---
title: "Unit2_Module2"
author: "Sushil Paudel"
date: "2024-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
# Import necessary libraries
l <- c("geospaar", "knitr")
lapply(l, require, character.only = TRUE)
```

## 1. 

Create a subset of districts by extracting districts 22, 26, 53, and 54. 
Call it districts_ss. Use the extent of districts_ss (ext(districts_ss)) to 
define the extent of a new raster r, which should have a resolution of 0.1°. 
Use r as a template for creating two new rasters, rsamp and rrandn. rsamp should
be filled with randomly selected integers ranging between 10 and 50. rrandn 
should be filled with random numbers drawn from a normal distribution (rnorm) 
that has a mean of 30 and standard deviation of 5. Use a seed of 1 in set.seed. 
Stack rsamp and rrandn (name the stack s), mask that by districts_ss, and
plot's using plot_noaxes.

```{r, warning = FALSE, Message = FALSE}
set.seed(1)

# add the districts data set
districts <- st_read(system.file("extdata/districts.geojson", 
                                 package = "geospaar"))


# extract the districts 22, 26, 53, 54
districts_ss <- districts %>% slice(c(22, 26, 53, 54))
kable(districts_ss)


# create a rasters templetes
r1 <- rast(ext(districts_ss), 
              resolution = 0.1, crs = districts)

r2 <- rast(ext(districts_ss), 
                     resolution = 0.1, crs = districts)


# Assign the values for each raster templates
values(r1) <- sample(10:50, size = ncell(r1), replace = TRUE)
values(r2) <- rnorm(ncell(r2), mean = 30, sd = 5)

# assign the names
names(r1) <- "rsamp"
names(r2) <-  "rrandn.rsamp"

# write a rasters
writeRaster(r1, overwrite = TRUE, 
            filename = file.path(tempdir(), "rsamp.tif"))

writeRaster(r2, overwrite = TRUE, 
            file.path(tempdir(), "rrandn.rsamp.tif"))


# read rasters 
rsamp <- rast(file.path(tempdir(), "rsamp.tif"))
rrandn.rsamp <- rast(file.path(tempdir(), "rrandn.rsamp.tif"))



#Stack rsamp and rrandn rasters and mask by districts_ss
s <- rast(list(rsamp, rrandn.rsamp)) %>% 
  mask(x = ., mask = districts_ss)

# plot stacked raster
par(mar = c(0, 0, 0, 0))
plot_noaxes(s)


```

## 2.
Disaggregate s[[1]] to a resolution of 0.025°, using bilinear interpolation, 
calling the result s2_1d. Select all areas of s2_1d that have values > 35, 
creating a new raster s2_1gt35. Set the values of s2_1gt35 that equal 0 to NA. 
Then convert the resulting raster into an sf object called s2poly. 
Plot the resulting polygons over s2_1d.


```{r}
# Disaggregate s[[1]] to a resolution of 0.025°
disagg(s[[1]], fact = 4, method = 'bilinear') -> s2_1d

# s2_1d that have values > 35
s2_1gt35 <- mask(s2_1d, s2_1d > 35)

# Set the values of s2_1gt35 that equal 0 to NA.
val <-  s2_1gt35 == 0
s2_1gt35[val] <- NA

# convert the resulting raster into an sf object called s2poly.
s2poly <- as.polygons(s2_1gt35, dissolve = TRUE) %>% st_as_sf 

# Plot the resulting polygons over s2_1d.
par(mar = c(0, 0, 0, 0))
plot(s2_1d)
plot(st_geometry(s2poly), add = TRUE)



```

## 3.
Create a new grid from the extent of districts that has a resolution of 0.5° 
(call it zamr), assigning all cells a value of 1. Then recreate the farmersr 
dataset–a raster that sums the number of farmers falling within each grid cell. 
Mask the results using districts, and then 
plot the result onto a grey background of Zambia. 

```{r, message = FALSE, error = FALSE}
# Create new grid from the extent of districts 
zamr <- rast(ext(districts), resolution = 0.5, crs = districts)
values(zamr) <- 1
names(zamr) <- "zamr" 

# create a raster that sums the number of farmers falling within each grid cell.
farmers <- read_csv(system.file("extdata/farmer_spatial.csv", 
                                  package = "geospaar"))
farmers %>% 
  distinct(uuid, .keep_all = TRUE) %>% 
  dplyr::select(x, y) %>% 
  mutate(count = 1) %>% 
  st_as_sf(., coords = c("x", "y")) %>% 
  st_set_crs(4326) %>% 
  st_transform(st_crs(districts)) -> farmers


farmersr <- rasterize(x = farmers, y = zamr, field = "count", fun = sum) %>% 
  terra::mask(., districts)

# plot the result onto a grey background of Zambia.
par(mar = c(1, 0, 0, 0))
plot(st_geometry(st_union(districts)), col = "grey",
     main = expression(paste("N farmers per 0.5", degree, " cell")))
plot(farmersr, add = TRUE, ext = districts)

```

## 4.
Convert the rasterized farmers counts (farmersr) back into an sf points object
farmersrpts. Create a new version of zamr at 0.05°, and then calculate the 
distance between these points and every other location in Zambia, creating an 
output grid of distances, called dist_to_farmers, which you mask by districts. 
Plot dist_to_farmers in kilometers (i.e. divide it by 1000) using plot_no_axes, 
with farmersrpts overlaid as black solid circles. 

```{r, rmessage = FALSE, error = FALSE}
# Conmvert the rasterized farmers counts into an sf points object
farmersrpts <- as.points(x = farmersr) %>% 
  st_as_sf
# create new version of zamr
zamr <- rast(ext(districts), resolution = 0.05, crs = districts) %>% 
  distance(x = ., y = farmersrpts) %>% 
  mask(., districts) -> dist_to_farmers

# plot
par(mar = c(0, 0, 0, 0))
plot_noaxes(dist_to_farmers/1000)
plot(farmersrpts, add = TRUE, col = "black", pch = 20)


```

## 5. 
Use geodata’s worldclim_country function to grab WorldClim’s mean temperature 
(“tavg”) dataset at a resolution of 2.5 (note this is not degrees, but minutes 
of a degree), and download it to somewhere on your local disk. That will give a
SpatRaster with 12 layers, with each layer representing the average monthly 
temperature for each grid cell on the planet. Calculate the annual mean 
temperature for each cell, and then mask the result using districts to get your
final raster, zamtmean. Plot the result.
```{r, rmessage = FALSE, error = FALSE}
# download the worldclim_country mean temperature data
tavg <- geodata::worldclim_country(country = "Zambia",
                                   res = 2.5,
                                   var = "tavg", 
                                   path = file.path(tempdir()),
                                   version = "2.1") %>% 
  app(., mean) %>% 
  mask(., districts) -> zamtmean


# Plot mean annual temperature
par(mar = c(0, 0, 0, 0))
plot(zamtmean, 
     main = expression(paste("Annual Mean Temperature of Zambia in ", degree, "C")))
plot(st_geometry(districts), add = TRUE)


```

## 6. 
Classify the temperature data into three categories, low, medium, and high, 
using <20°, 20-24°, and >24° as the break points for determining the classes. 

#### **What are the functions floor and ceiling doing?**

*The floor() returns integer as a lower bounds where as* 
*ceiling() returns integer by rounding up to the upper digit as an integer*
*For instance floor(trng[1]) returns 15 instead of 15.04167. Similarly,*
*ceiling(trng[2]) returns 28 instead of 27.19167.*

Call the reclassified temperature raster zamtclass.


Make the map 
using plot_noaxes with a categorical legend, and using the colors “blue”, 
“yellow2”, and “red” for the three classes.


```{r, rmessage = FALSE, error = FALSE, fig.cap = "Classified Temperature of Zambia" } 
trng <- global(zamtmean, range, na.rm = TRUE)
reclmat <- cbind(c(floor(trng[1]), 20, 24), 
                 c(20, 24, ceiling(trng[2])), 
                 1:3) %>% as.matrix.data.frame()


zamtclass <- terra::classify(x = zamtmean, 
                             rcl = reclmat, include.lowest = TRUE)
par(mar = c(0, 0, 0, 0))
plot_noaxes(zamtclass,
            col = c("blue", "yellow2", "red"),
            plg = list("top", 
            legend = c("Low", "Medium", "High"),
            fill = c("blue", "yellow2", "red"),
            title = "Classified Temperature"))

plot(st_geometry(districts), add = TRUE)



```

## 7. 
Recreate the zamprec dataset (chunk 48), then calculate the mean precipitation
within each temperature zone defined by zamtclass. Call the resulting matrix z.
Map the mean zonal precipitation values in z onto each temperature zone (using 
the subst function with zamtclass as the target; remember that zonal returns a 
matrix, and that subst requires equal length vector for the “from” and “to” 
values. Call the new raster zamprecz, and then plot it using plot_noaxes, with 
a custom legend (as done in Task 6), using the rounded zonal mean values 
(rounded) as the legend labels (legend = round(z$mean)). Use colors “yellow2”, 
“green3”, and “blue” for the three classes.

```{r, rmessage = FALSE, error = FALSE}
wcprec <- geodata::worldclim_country(var = "prec", res = 2.5, 
                                     country = "Zambia", 
                                     path = tempdir())

prec <- (app(wcprec, mean) %>% mask(., districts))

z <- zonal(x = prec, z = zamtclass, fun = "mean", na.rm = TRUE)
  
subsmat <- z %>% dplyr::select(1:2)
zamprecz <- subst(x = zamtclass, from = subsmat[, 1], to = subsmat[, 2])


par(mar = c(1, 0, 0, 0))
plot_noaxes(zamprecz, main = "Mean Annual Precipitaion",
            col = c("yellow2", "green3", "blue"),
            plg = list("top", 
            legend = round(z$mean),
            fill = c("yellow2", "green3", "blue"),
            title = "Mean annual precipitation"))

plot(st_geometry(districts), add = TRUE)

```

## 8.
Use geodata::elevation_30s again to download the elevation raster for Zambia 
(call it dem). Aggregate it to the same resolution as chirps, using the default 
mean aggregation, and mask it using districts. Call that dem5. Use terrain to 
calculate aspect from dem5 (call it aspect), selecting degrees as the output 
value. 
Then find all west-facing aspects (aspects >247.5 and <292.5), and all 
east facing aspects (>67.5 and <112.5), making new rasters respectively named 
west and east, e.g. west <- aspect > 247.5 & aspect < 292.5). 
Stack these together with aspect and make a three-panel plot with plot_noaxes 
with titles “Aspect”, “West”, and “East”.

```{r, message = FALSE, error = FALSE, fig.align='center'}
dem <- geodata::elevation_30s(country = "Zambia", path = file.path(tempdir()))
chirps <- terra::rast(system.file("extdata/chirps.tif", package = "geospaar"))

# resampling the resolution as chirps
dem5 <- terra::aggregate(x = dem, fact = 6, fun = mean) %>% 
  mask(., districts) %>% 
  terrain(., v = "aspect", unit = "degrees")-> aspect


west <- aspect > 247.5 & aspect < 292.5
east <- aspect >67.5 & aspect < 112.5

stack_r <- c(aspect, west, east)
names(stack_r) <- c("Aspect", "West", "East")

par(mfrow = c(1, 3))
plot_noaxes(stack_r[[1]], plg = list(legend = c("Aspect"),
            title = "Aspect (degrees)"), main = "Aspect")

plot_noaxes(stack_r[[2]], col = c("grey","seagreen"),
            plg = list(legend = c("Others", "West"),
                                     fill = c("grey","seagreen"),
                                     title = "Aspect"), main = "West")

plot_noaxes(stack_r[[3]], col = c("grey", "lightseagreen"), 
            plg = list(legend = c("Others", "East"),
                                     fill = c("grey","lightseagreen"),
                                     title = "Aspect"), main = "East")
```

## 9.
Using a random seed of 1, create two random samples of 100 each. The first one 
should be collected from within the west-facing cells (i.e. only be drawn from
cells in west that have a cell of one), and the second from east-facing cells.
To do this, set the cells equal to 0 in east and west to NA 
(e.g. west[west == 0] <- NA). Once you have collected those, convert the 
resulting objects to sf, and use those two sets of points to extract temperature
values from zamtmean into a tibble temp_stats, 
Then use temp_stats with ggplot to make side-by-side boxplots to compare the 
distributions of west and east facing temperatures, modeled on the example in 
Chunk 40.
```{r, rmessage = FALSE, error = FALSE}
set.seed(1)

# set the cells equal to 0 in east and west to NA
west[west == 0] <- NA
east[east == 0] <- NA

# random sample point within east and west facing cells.
rs1 <- spatSample(x = west, size = 100, method="random", 
                  xy = TRUE, na.rm = TRUE) %>% 
  st_as_sf(., coords = c("x", "y"), crs = st_crs(west)) -> westpts

westpts

rs2 <- spatSample(x = east, size = 100, method="random", 
                  xy = TRUE, na.rm = TRUE) %>% 
  st_as_sf(., coords = c("x", "y"), crs = st_crs(east)) -> eastpts

eastpts



# values from zamtmean into a tibble temp_stats
temp_stats <- bind_rows(tibble(temp = terra::extract(zamtmean, 
                                                  westpts)$mean, dat = "West"), 
                        tibble(temp = terra::extract(zamtmean,
                                                  eastpts)$mean, dat = "East"))

temp_stats


# theme template
bp_theme <- theme(legend.title = element_blank(), 
                  axis.text.x = element_blank(),
                  axis.ticks.x = element_blank(), 
                  panel.grid.major.x = element_blank(), 
                  panel.grid.minor.x = element_blank(), 
                  panel.background = element_rect(fill = "grey95"))


# plot
par(mar = c(0, 0, 0, 0))
ggplot(temp_stats) +
  geom_boxplot(mapping = aes(y = temp, fill = dat), position = "dodge2") +
  scale_fill_manual(values = c("lightseagreen", "seagreen"))+
  ggtitle("Mean Annual Temperature") + xlab(NULL) + ylab("Degree Celcious") +
  bp_theme
  


```

## 10. 
Extract the centroids from each district in districts (call it dcent), and 
reproject the points to Albers, using the st_crs(roads). 
Reproject zamtmean to Albers also, making the new resolution (5 km, i.e. 5000 m), 
using bilinear interpolation (call it zamtmeanalb). 

Then use dcent to extract the temperature values from zamtmeanalb (add the 
values to dcent as a new variable “temp” using mutate). Use gstat to create an 
IDW model (call it idw). To make the IDW work,
which isn’t sf compliant, some extra work will be required, as shown below 
(this is the step needed after the extract of temperature values)

`dcent <- bind_cols(`
  `dcent %>% data.frame %>% dplyr::select(-geometry) %>% as_tibble, `
  
  `st_coordinates(dcent) %>% as_tibble) %>% rename(x = X, y = Y)`
  
This yields a tibble with columns x and y that are needed by gstat. After 
running gstat, map the interpolated temperatures using zamtmeanalb as a target 
object (it won’t be overwritten) and idw as the model. Make sure you mask the 
result to the boundaries of Zambia, using zamtmeanalb as the mask. Call the new 
interpolated, masked grid zamtidw. Plot the result side by side with zamtmeanalb
for comparison using plot_noaxes with titles 
“Real Temperature” and “IDW Temperature”.

```{r, message = FALSE, error = FALSE}
roads <- st_read(system.file("extdata/roads.geojson", package = "geospaar"))
```

```{r, rmessage = FALSE, error = FALSE}
library(sp)

# Extract the centroids from each district and reproject
districts %>% st_centroid %>% 
  st_transform(crs = st_crs(roads)) -> dcent


# reproject zamtmean to Albers
zamtmeanalb <- project(x = zamtmean, y = roads, 
                       res = 5000, method = "bilinear")


# Extract values from zamtmean using dcent
dcent %>% mutate(temp = terra::extract(zamtmeanalb, .)$mean) -> dcent


# Create an empty grid where n is the total number of cells
dcent <- bind_cols(dcent %>% 
                     data.frame %>% 
                     dplyr::select(-geometry) %>% 
                     as_tibble, st_coordinates(dcent) %>% 
                     as_tibble) %>% rename(x = X, y = Y)

rgrid <- as.data.frame(zamtmeanalb)

# idw <- gstat::idw(temp ~ 1, dcent, newdata = rgrid, idp = 2.0)

# zamtidw <- rast(idw)

par(mfrow = c(1, 3))
plot_noaxes(zamtmeanalb, main = "Real Temperature")
# plot_noaxes(zamtidw, main = "IDW Temperature")


```







