---
title: "unit2-module1"
author: "Sushil Paudel"
date: "2024-03-20"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{unit2-module1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing necessary libraries to perform the vignette tasks.

```{r}
l <- c("sf", "dplyr", "tidyr", "ggplot2", "geospaar")
lapply(l, require, character.only = TRUE)

```

**6.2.1**
Read **farmers_spatial.csv**, **districts.geojson**, and **roads.geojson**
datasets. 

Reduce the size of the farmers data by first selecting distinct observations by
uuid, x, y, season, i.e. use distinct(uuid, x, y, season). 

After that convert it to an sf object.

Reproject the farmers and districts data to Albers projection 
(using the CRS from roads), naming each farmers_alb and districts_alb. 

Ideally (worth an extra 0.5 points) you will do all the necessary
steps to create farmers_alb and districts_alb in one pipeline.

```{r}

# reading farmers_spatial.csv
farmers <- read_csv(system.file("extdata/farmer_spatial.csv", 
                                  package = "geospaar"))
farmers

# reading districts.geojson 
districts <- st_read(system.file("extdata/districts.geojson", 
                                 package = "geospaar"))

districts

st_crs(districts)
# readidistricts# reading roads.geojson
roads <- st_read(system.file("extdata/roads.geojson", package = "geospaar"))
roads


# Reducing the size of the farmers data
farmers <- farmers %>% 
  distinct(uuid, x, y, season)
farmers  

# Converting to sf object
farmers <- st_as_sf(farmers, coords = c("x", "y"))
farmers

# Reprojecting farmers and districts crs to that of roads datasets.
farmers_alb <-  farmers %>% 
  st_set_crs(4326) %>% 
  st_transform(crs = st_crs(roads))
farmers_alb

districts_alb <- districts %>% 
  st_set_crs(4326) %>% 
  st_transform(crs = st_crs(roads))
districts_alb

st_crs(farmers_alb)
st_crs(districts_alb)

```

**6.2.2**
Create a plot using sf::plot that shows all three datasets on one map, with 
districts_alb in grey, with roads in red over that, and farmers_alb as a blue 
cross over that. Use the relevant chunk arguments to center the figure in the 
vignette html, and to have a height of 4 inches and a width of 6 inches. 
The figure should have 0 margins all around.

```{r fig.width = 6, fig.height = 4, fig.align = "center"}
#Setting margin(bottom, left, top, right)
par(mar = c(0, 0, 0, 0))
# plotting districts dataset
plot(districts_alb %>% st_geometry(), col = "grey") 
# Plotting roads dataset
plot(roads %>% st_geometry(), add = TRUE, col = "red")
# plotting farmers dataset
plot(farmers_alb %>% st_geometry(), add = TRUE, col = "blue", pch = 20)


# title("The map showing roads and farmers over the district map")
legend(x = "bottomright", 
       pch = c(20, 1, 15), 
       col = c("blue", "red", "grey"), 
       legend = c("farmers","roads", "districts"),
       bty = "n")
```

**6.2.3**
Make the same plot above using ggplot and geom_sf. 

When adding farmers_alb to the plot, use pch = "+" and 
size = 3 as arguments to geom_sf. 

Add the function theme_bw() to the ggplot construction chain,
to get rid of the grey background.
Make the “fill” (rather than “color”) of districts_alb grey. 

Center the figure using chunk options and make the figure width 5 inches 
and height 6 inches.

```{r fig.width = 5, fig.height = 6, fig.align = "center"}

ggplot() +
  geom_sf(data = districts_alb, fill = "grey") +
  geom_sf(data = roads, col = "red") +
  geom_sf(data = farmers_alb, pch = "+", size = 3, col = "blue") +
  theme_bw()
 
```

**6.2.4**
Select from districts_alb the district representing the 50th percentile area, 
i.e. the median area, and save that district into a new object median_dist. 
Plot it in “khaki” on top of grey districts_alb. Give the plot a title 
“The median area district”. Same plot dimensions in the vignette html as for 
Task 2, but a leave a space of 1 at the top in the plot mar.

```{r fig.width = 6, fig.height = 4, fig.align = "center"}

pertile <- function(x) quantile(x, probs = c(0, 0.5))

median_dist <- districts_alb %>% 
  mutate(area = as.numeric(units::set_units(st_area(.), "km^2"))) %>% 
  filter(area == pertile(area)) 

median_dist

#Setting margin(bottom, left, top, right)
par(mar = c(0, 0, 1, 0))
plot(st_geometry(districts_alb), col = "grey", reset = FALSE)
plot(st_geometry(median_dist), add = TRUE, col = "khaki")

title("The median area district")
legend(x = "bottomright", 
       pch = 15, 
       col = c("khaki", "grey"), 
       legend = c("median_dist","district_alb"),
       bty = "n")


```

**6.2.5**
Convert the median_dist to its centroid point. Call it median_distp. filter the 
farmers_alb data for season 1, and then find the 20 closest season 1 farmers to
median_distp. To do that, create the new object closest_20farmers by using 
mutate with st_distance to create a new variable dist (convert it to numeric),
and then arrange by variable dist and slice out the top 20 observations. Plot
districts_alb in grey, median_dist over that in khaki, median_distp as a solid 
purple circle, farmers_alb in blue, and closest_20farmers in red. Zero margins 
and width of 6 inches and height of 4 inches.

```{r  fig.width = 6, fig.height = 4, fig.align = "center"}
##Finding the centroid of a median district
median_distp <- median_dist %>% 
  st_as_sf() %>% 
  st_centroid()
median_distp

## Filtering the farmers of season 1
farmers_s1 <- filter(farmers_alb, season == 1)
farmers_s1

## Finding the 20 farmers of season 1 closest to the centroid
closest_20farmers <- mutate(farmers_s1, 
                            dist = st_distance(geometry, median_distp) %>% 
                              as.numeric()) %>% arrange(dist) %>% slice(1:20)
closest_20farmers

## Plotting the data
par(mar= c(0, 0, 0, 0))
ggplot() + geom_sf() +
  geom_sf(data = districts_alb, fill = "grey", col = "black") +
  geom_sf(data = farmers_alb, col = "blue") +
  geom_sf(data = closest_20farmers, col = "red")+
  geom_sf(data = median_distp, col = "purple", size = 3)+
  geom_sf(data = median_dist, fill = "khaki") +
  ggtitle ("Districts with 20 closest farmers")+
  theme_bw()

```


**6.2.6**
Create a rectangular sf polygon called mypol using the x coordinates 30 and 31 
(minimum and maximum) and y coordinates -10 and -11. 
Assign it crs = 4326 and transform it to Albers. 

Select from districts_alb the districts that intersect mypol, and 
plot in “grey40” over districts_alb in grey, and 
plot over that mypol without any fill but just a yellow border. 

Calculate the area in ha of mypol and report it in your vignette below this plot. 

Zero margins and width of 6 inches and height of 4 inches.

```{r fig.width = 6, fig.height = 4, fig.align = "center"}
coords <- cbind("x" = c(30, 30, 31, 31, 30),
                "y" = c(-10, -11, -11, -10 ,-10))
coords
mypol <- st_polygon(x = list(coords)) %>% 
  st_sfc %>% 
  st_sf(ID = 1, crs = 4326) %>% 
  st_transform(crs = st_crs(roads))

mypol <- mypol %>% 
  mutate(area_ha = as.numeric(units::set_units(st_area(.), "ha")))


dists_int <- districts_alb %>% 
  slice(st_intersects(x = mypol, y = districts_alb)[[1]])  
dists_int


par(mar = c(0, 0, 0, 0))
plot(st_geometry(districts_alb), col = "grey")
plot(st_geometry(dists_int), col = "grey40", add = TRUE)
plot(st_geometry(mypol), border = "yellow", add = TRUE)

area <- round(mypol$area_ha,2)
paste0("The area of mypol rectangular feature in hectare is ", {area},".")

```


**6.2.7**
Create mypol_dist_int from the intersection of mypol and districts_alb, 
recasting the intersected districts to multipolygons, and 
adding an area variable onto it that reports areas of intersections in hectares. 
Do all that in one pipeline. 

Plot mypol_dist_int in rainbow colors over districts_alb. 
Zero margins and width of 6 inches and height of 4 inches. 

Report the mean and median of interections in the vignette below the plot.

```{r fig.align = 'center', fig.height = 4, fig.width = 6}
mypol_dist_int <- st_intersection(x = mypol, y = districts_alb) %>%
  mutate(area_ha = as.numeric(units::set_units(st_area(.), "ha")))

mypol_dist_int


par(mar = c(0, 0, 0, 0))
rbcols <- rainbow(n = nrow(mypol_dist_int))
plot(st_geometry(districts_alb), col = "grey")
plot(st_geometry(mypol_dist_int), col = rbcols, add = TRUE)

area_report <- data.frame(distName = mypol_dist_int$distName, 
           area_ha = round(mypol_dist_int$area,2))
print("The detail of intersection is shown in table below:")
print(area_report)


muarea <- round(mean(mypol_dist_int$area_ha),2)
medarea <- round(median(mypol_dist_int$area_ha),2)

paste("The mean of intersections is", {muarea},"ha.")
paste("The median of intersections is",{medarea},"ha.")



```


**6.2.8**
Find the shortest and longest roads in Zambia, and place the selected roads into
a new object (roads_extreme). To do this, you will need to arrange roads by 
length and then slice to get the first and last observations (of course you need
to first calculate length). Do that as one pipeline. Then calculate a 50 km 
buffer around those two roads (roads_extreme_buff). Plot roads_extreme_buff in 
blue over districts_alb in grey, and add roads_extreme on top of that as red 
lines (use lwd = 3 in the plot). Zero margins and width of 6 inches and height 
of 4 inches.

```{r fig.height=4, fig.width=6, fig.align='center'}
roads_extreme <- roads %>% 
  mutate(length_km = as.numeric(units::set_units(st_length(.),"km"))) %>% 
  arrange(length_km) %>% 
  slice(c(1, n()))  
roads_extreme  

roads_extreme_buff <- st_buffer(roads_extreme, dist = 50000)
roads_extreme_buff

par(mar = c(0, 0, 0, 0))
plot(st_geometry(districts_alb), col = "grey")
plot(st_geometry(roads_extreme_buff), add = TRUE, col = "blue" )
plot(st_geometry(roads_extreme), add = TRUE, col = "red", lwd = 3)

```

**6.2.9**
Select a random sample of 10 points in the smallest object in roads_extreme_buff
, and one of 50 in the largest object. Use a single call to st_sample to do 
that. Use a seed of 2. Plot those points as yellow solid points over the same
map created in Task 8 above. Use the same dimensions.

```{r fig.height=4, fig.width=6, fig.align='center'}
set.seed(2)
sample <- roads_extreme_buff %>% 
  mutate(area_sqkm = as.numeric(units::set_units(st_area(.), "km^2"))) %>% 
  arrange(area_sqkm) %>% 
  st_sample(., size = c(10, 50))

sample

par(mar = c(0, 0, 0, 0))
plot(st_geometry(districts_alb), col = "grey")
plot(st_geometry(roads_extreme_buff), add = TRUE, col = "blue" )
plot(st_geometry(roads_extreme), add = TRUE, col = "red", lwd = 3)
plot(st_geometry(sample), add = TRUE, pch = 20, col = "yellow")

```

**6.2.10**
Your final task is to intersect roads with the buffer of the longest road in 
roads_extreme_buff (roads_int). Plot the buffer of the longest road in blue over
the districts in grey, and then roads_int as red lines. Use the same dimensions 
as the previous two plots. Report the total distance of intersected roads in km 
in the vignette below the plot.

```{r fig.height=4, fig.width=6, fig.align='center'}
roads_int <- roads %>% 
  st_intersection(x = roads_extreme_buff, y = .) %>% 
  mutate(length_km = as.numeric(units::set_units(st_length(.), "km")))

roads_int

par(mar = c(0, 0, 0, 0))
plot(st_geometry(districts_alb), col = "grey")
plot(st_geometry(roads_extreme_buff), add = TRUE, col = "blue" )
plot(st_geometry(roads_int), add = TRUE, col = "red", lwd = 3)


roads_dist <- sum(st_length(roads_int))
cat("The total distance of intersected roads in km is", {roads_dist})
```

